---
title: "Group Project 2"
subtitle: "Biology 368/664 Bucknell University"
output: html_notebook
author: Prof. Ken Field
date: 19 Feb 2019
---

## Objectives

1. Critically read a scientific paper and assess the evidence provided
2. Assess a published paper for issues with reproducibility and transparency
3. Identify questionable research practices and describe appropriate methods to prevent them
4. Complete a preregistration plan describing a confirmatory study following the principles of 

## Groups

We will work in pairs for this project, with each pair being different from the Project 1 groups.

```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("knitr")) install.packages("knitr"); library(knitr)
```

```{r echo=FALSE, results='asis'}
Project2 <- read.csv("../../Biol364_664roster.csv", stringsAsFactors=FALSE)
Project2$Group <- c("C",
                    "A",
                    "D",
                    "B",
                    "F",
                    "C",
                    "A",
                    "D",
                    "E",
                    "B",
                    "F",
                    "E")
table(Project2$Group)
kable(arrange(Project2, Group))
```

## Choosing Papers

Each group will choose a paper in lab on Tuesday, Feb 19. You may want to browse the papers available and discuss them with you group partner prior to lab.

```{r}
dir("papers", pattern = "pdf", full.names = TRUE, ignore.case = TRUE)
```

## Grading

Each student will be expected to complete the following tasks to earn 85% of the points available for this assignment (21/25).

- Read and discuss your paper with your group partner
  + Focus on the methods and results
  + For each conclusion evaluate the strength of the supporting evidence

#Methods and results (basic):
-Started with 88,000 chemicals put into wells with C. elegans
-115 statistically extended life 
-They chose one that that extended life by 20%-- why? There were some that extended it more. -This one is structurally related to compounds used as antidepressants in humans. Specifically impacts serotinin signialing.
-Chose 20 compounds that influence serotonin signaling. Of these 4 increased lifespan by 20-33%(is this significant)? These are a lot like 2 drugs Miaserin (used for humans) and methiotherpin. 
-Investigated these further. 
Table 1. looks important- I'm a little confused by number of experiments and number of animals 
  
  
- 1) Complete the nr-reporting-summary.pdf checklist (open in Adobe Acrobat) 
  + For any deficient items on the checklist, describe the information needed
  + Refer to nr-reporting-life-sciences-research.pdf for guidelines
KF: Good work on this. The explanations were very helpful (although tiny!)
The only missing information that I noticed was about the strain of C. elegans used. Was there any info on this in the paper?
  
2 - Determine if you could repeat the analysis that is most essential to conclusions of the paper
  + If data or methods are not available, look for follow-up studies that provide the required details
  
The analysis they used to determine the signifigance of the effects of the drugs they tested is called the log rank test or Mantel-Hantzael test. This test is well designed for asssessing survivorship because it allows one to make comparisons of the number of binary events between a treated and control group at multiple time points. In the case of survivorship, the binary evet is death and thus it can account for differences in the number of deaths between two groups at every time that the number of surviving subjects are counted. This treats every time point like an independent sample and allows an  investigator to compare and weight the differences of two groups each time there are events and then sum their contributions to an overall difference in a similar manner to the chi squared analysis. 

In order to complete this analysis one needs to essentailly construct two by two tables at every time point:
       Control   Treatment   
Alive    x           x

Dead     y           y

In this paper the data they report only includes the mean lifespan in days rather than the breakdown of the number of death events and surviving c elegans for each day that they were measured. Since they only report these summary statistics along with the corresponding p-values, it is imposible to replicate their Mantel-Haentzel analysis that yielded those p-values.

https://www.biochemia-medica.com/en/journal/24/1/10.11613/BM.2014.003
http://www.math.ucsd.edu/~rxu/math284/slect3.pdf
https://www.youtube.com/watch?v=YjU2sFHeaBE

Although there are additional studies from this lab that use the same c-elegans survivorship paradigm and use the same mantel haetzel test, they do not prpovide the missing raw data from this study required to perform the analysis independently. 
  + If follow-up studies are not available, describe the missing information needed to perform the analysis

The major missing information is the breakdown of the death events and survivorship by day for each drug they tested as well as the complete screen of compounds they tested as opposed to just including the serotonin related tests. 

http://www.biostathandbook.com/cmh.html
   
KF: Nice explanation of the test used and the reasoning behind it.  
  
- 3) For each Questionable Research Practice discussed in Fraser *et al*:
  + Find evidence to exclude this QRP, or
  + Describe steps that should have been taken to prevent this QRP (see Forstmeier *et al*)
  
  
#Initial thoughts

The biggest question I have regarding the research is centered around this statement from the online materials: 
Statistical analysis
STATA8 software was used for analysis. Comparisons and P value calculations were made between treated and untreated animals of the same strain using the log-rank test (Mantel–Haenszel). We observed the death of 98.6% of the animals (excluding screen). Animals that were still alive at the end of an experiment (1.4%) were analysed as alive up to this point with unknown time of death (censoring). Wells containing more than 19 animals were excluded from the analysis.

It is clear looking at the supplementary materials that most treatments had less individuals than in the treatment group than the control. Was this due to death of those in the treatment?

I am also curious about pseudo-replication in well plate. 
KF: That is an interesting question. 

Also of concern is if standard methods were used.
eg  "The fraction of animals alive was scored on the basis of body movement." Is this typical for C. elegans?
KF: Do you think that this might be subjective? If so, it would be important to blind the experimenters to the treatment, right?

- 3) For each Questionable Research Practice discussed in Fraser *et al*:
  + Find evidence to exclude this QRP, or
  + Describe steps that should have been taken to prevent this QRP (see Forstmeier *et al*)
  
KF: Good work on this section.  

1- unreported values
They did not report any specifics on the 88,000 compounds they tested in their initial screen. Even more concerning was their lack of information on the selection criteria for the compounds they decided to retest in larger populations beyond the generic statement of "On the basis of the fraction of live animals per well relative to controls".They also failed to report anything on what the other 113 compounds were that significantly increased lifespan and instead only showed results for serotonin related treatments and moved forward with that despite quoting that some compounds increased lifespan by a large margin more than serotonin related treatments.

They should have provided data on the specifics of each treatment they screened with. 

2- unreported covariates
Although it seems unlikely, since their readout for survivial was movement it could be possible that some of their drug treatments could have affected movement and therefore the change in lifespan associated with the drugs may be realted more to movement than to actual changes in longevity. 

They should have assessed the effect sof these drugs on movement in a different test. 

3- harking
Yes... used many chemicals and it just so happened that the antidepressant showed interesting results. They were working wiht a screen of close to 88,000 compounds and it seems unlikely that they had a hypothesis of the importance of serotonin related treatments prior to such a screen. 


4- unreported models
N/A they worked exclusively in c-elegans and reported on their survivorship model

5- rounding ps
It is possible that they rounded p-values as the in depth report for each serotonin related trial only shows p-values to two significant figures an dhtere are p-values of exactly .05 for some treatments. 

They should have reported p-values to more decimal places

6- excluding data
Concern that they excluded data from nematodes that died quickly. SHould have been more transparent about their protocol for deciding exactly when a trail started/ended. 

7- adding data
Number of experiments was not constant, unsure if they added more experiments as needed? Solved by being transparent about methods. Eg we attempted to use 150 nematodes for each treatment, any individuals that dies before "adulthood" were not used in the assay."

8- switching analyses
N/A only used one analysis that they reported:log-rank test Mantel–Haenszel test

9- undisclosed problems
Potentially... such as young nematodes dying. 

 
10- fabrication
Unlikely...


- 4) Design a follow-up study to replicate the most crucial experiment
  + Complete the PreregChallengeTemplate.Rmd form for this replication experiment
KF: This has just been started but a lot of work remains.  
  
- 5) Present summary of reproducibility assessment in class

Each group should use an **Acknowledgements** section to document the participation of each member and the collaboration within and between groups.

Additional credit will be awarded for providing assistance to other groups and for completion of the analysis that supports the most essential conclusion.

**Acknowledgements:** 
Cheyenne:
-nature reporting summary
-analysis research
-
Ian: 
?

Jordan and Matt for summary form and organization tips
