---
title: "Group Project 2"
subtitle: "Biology 368/664 Bucknell University"
output: html_notebook
author: Prof. Ken Field
date: 19 Feb 2019
---

## Objectives

1. Critically read a scientific paper and assess the evidence provided
2. Assess a published paper for issues with reproducibility and transparency
3. Identify questionable research practices and describe appropriate methods to prevent them
4. Complete a preregistration plan describing a confirmatory study following the principles of 

## Groups

We will work in pairs for this project, with each pair being different from the Project 1 groups.

```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("knitr")) install.packages("knitr"); library(knitr)
```

```{r echo=FALSE, results='asis'}
Project2 <- read.csv("../../Biol364_664roster.csv", stringsAsFactors=FALSE)
Project2$Group <- c("C",
                    "A",
                    "D",
                    "B",
                    "F",
                    "C",
                    "A",
                    "D",
                    "E",
                    "B",
                    "F",
                    "E")
table(Project2$Group)
kable(arrange(Project2, Group))
```

## Choosing Papers

Each group will choose a paper in lab on Tuesday, Feb 19. You may want to browse the papers available and discuss them with you group partner prior to lab.

```{r}
dir("papers", pattern = "pdf", full.names = TRUE, ignore.case = TRUE)
```

## Grading

Each student will be expected to complete the following tasks to earn 85% of the points available for this assignment (21/25).

- Read and discuss your paper with your group partner
  + Focus on the methods and results
  + For each conclusion evaluate the strength of the supporting evidence

#Methods and results (basic):
-Started with 88,000 chemicals put into wells with C. elegans
-115 statistically extended life 
-They chose one that that extended life by 20%-- why? There were some that extended it more. -This one is structurally related to compounds used as antidepressants in humans. Specifically impacts serotinin signialing.
-Chose 20 compounds that influence serotonin signaling. Of these 4 increased lifespan by 20-33%(is this significant)? These are a lot like 2 drugs Miaserin (used for humans) and methiotherpin. 
-Investigated these further. 
Table 1. looks important- I'm a little confused by number of experiments and number of animals 
  
  
- 1) Complete the nr-reporting-summary.pdf checklist (open in Adobe Acrobat) 
  + For any deficient items on the checklist, describe the information needed
  + Refer to nr-reporting-life-sciences-research.pdf for guidelines
  
  
- 2) Determine if you could repeat the analysis that is most essential to conclusions of the paper
  + If data or methods are not available, look for follow-up studies that provide the required details
  + If follow-up studies are not available, describe the missing information needed to perform the analysis

#It is likely these analyese could be repeated since they only used one test: the log-rank test Mantel–Haenszel test
http://www.biostathandbook.com/cmh.html
However, since they have so many treatments I'm not positive this was the best test, though I'm not a biostats expert.   
  
- 3) For each Questionable Research Practice discussed in Fraser *et al*:
  + Find evidence to exclude this QRP, or
  + Describe steps that should have been taken to prevent this QRP (see Forstmeier *et al*)
  
#The biggest question I have regarding the research is centered around this startment from the online materials: 
Statistical analysis
STATA8 software was used for analysis. Comparisons and P value calculations were made between treated and untreated animals of the same strain using the log-rank test (Mantel–Haenszel). We observed the death of 98.6% of the animals (excluding screen). Animals that were still alive at the end of an experiment (1.4%) were analysed as alive up to this point with unknown time of death (censoring). Wells containing more than 19 animals were excluded from the analysis.

It is clear looking at the supplementary materials that most treatments had less individuals than in the treatment group than the control. Was this due to death of those in the treatment?

I am also curious about pseudo-replication in well plate. 

Also of concern is if standard methods were used.
eg  "The fraction of animals alive was scored on the basis of body movement." Is this typical for C. elegans?


- 3) For each Questionable Research Practice discussed in Fraser *et al*:
  + Find evidence to exclude this QRP, or
  + Describe steps that should have been taken to prevent this QRP (see Forstmeier *et al*)
1- unreported values
2- unreported covariates
3- harking
4- unreported models
5- rounding ps
6- excluding data
Concern that they excluded data from nematodes that died quickly
7- adding data
Number of experiments was not constant, unsure if they added more experiments as needed?
8- switching analyses
N/A only used one analysis
9- undisclosed problems
Potentially..
10- fabrication
Unlikely...


- 4) Design a follow-up study to replicate the most crucial experiment
  + Complete the PreregChallengeTemplate.Rmd form for this replication experiment
  
- 5) Present summary of reproducibility assessment in class

Each group should use an **Acknowledgements** section to document the participation of each member and the collaboration within and between groups.

Additional credit will be awarded for providing assistance to other groups and for completion of the analysis that supports the most essential conclusion.

**Acknowledgements:** 
Cheyenne:
-nature reporting summary
-analysis research
-
Ian: 

Jordan and Matt for summary form and organization tips
